Computing in the Field of Rationals

### 2012-06-26 rational_ext.tex
_

Say more: fractionals, entaglements, backtracking, superposition,
constrants, negative information. Negative types, backward information
flows, sets with negative types, iteration.. etc.

@@

We present the categorical structure with minimum commentary. More
details may be found in excellent references such as Barr and Wells
[CITE] and Selinger \cite{springerlink:10.1007/978-3-642-12821-94}.
For brevity, we don't re-state standard categorical definitions and
subsequently present their {{Pi}} equivalents. Instead, we directly
state categorical definitions in terms of {{Pi}} types and
combinators, and provide appropriate references to the former. For
example, associativity is denoted by the greek letter $\alpha$ (and
$\alpha^{-1}$) in standard definitions, whereas we use {{assocl*}}
(and {{assocr*}}).

@@

To show the right-to-left direction we proceed by induction of the
sequence of {{|-->*}} derivations. Again the case for primtive
isomorphisms follows trivially. Taking the case of {{c1+c2}} we are
required to show that given a sequence 
{{<c1+c2,v,C> |-->* {[c1+c2, v', C]} }} there is a derivation tree for
{{c1+c2 v |--> v'}}. In the case that {{v:v1+b2}} has the form
{{left~v1}}, this follows by observing that the given sequence must have the form
{{ <c1+c2, left v1, C> |--> <c1,v1,LeftP c2 C> |-->* {[c1,v1', LeftP c2 C]} |--> {[c1+c2,left v1',C]} }}. 
For the strictly smaller inner subsequence by induction we have
{{c1~v1 |--> v1'}} which lets us complete the derivation of
{{c1+c2~(left v1) |--> left v1'}}. The {{right v2}} follows similarly.

@@

For any execution path from {{v}} to {{v'}} in {{c}}, {{c{dagger} }}
has an execution path from {{v'}} to {{v}}.  Stated differently,
logical reversibility implies that {{c^{dagger} }} denotes the inverse
relation that {{c}} denotes, i.e.  {{c^^^={|(v',v)~|~(v,v') elem c|} }}.  
Intuitively logical reversibility holds because for every
{{eta*}} reduction in a world that transforms {{()}} into {{(1/v,v)}},
on reverse execution the adjoint {{eps*}} will transform {{(1/v,v)}}
back to {{()}}.

@@

Like {{Pi}}, {{PiEE*}} is a commutative bimonoidal catgeory.
Symmetric monoidal categories where every object had a dual object and
where dualizing preserves the monoidal tensor are referred to as
compact closed categories. Since divison is defined only for types of
non-zero size, not all objects have duals.  We show that {{PiEE*}}
respects compact closed structure for non-zero objects. We introduce
the term ``locally compact closed'' to refer to this notion.

>>>>

Say more: fractionals, entanglements, backtracking, superposition,
constraints, negative information. Negative types, backward information
flows, sets with negative types, iteration.. etc.

@@

We present the categorical structure with minimum commentary. More
details may be found in excellent references such as Barr and Wells
[CITE] and Selinger \cite{springerlink:10.1007/978-3-642-12821-94}.
For brevity, we don't re-state standard categorical definitions and
subsequently present their {{Pi}} equivalents. Instead, we directly
state categorical definitions in terms of {{Pi}} types and
combinators, and provide appropriate references to the former. For
example, associativity is denoted by the Greek letter $\alpha$ (and
$\alpha^{-1}$) in standard definitions, whereas we use {{assocl*}}
(and {{assocr*}}).

@@

To show the right-to-left direction we proceed by induction of the
sequence of {{|-->*}} derivations. Again the case for primitive
isomorphisms follows trivially. Taking the case of {{c1+c2}} we are
required to show that given a sequence 
{{<c1+c2,v,C> |-->* {[c1+c2, v', C]} }} there is a derivation tree for
{{c1+c2 v |--> v'}}. In the case that {{v:v1+b2}} has the form
{{left~v1}}, this follows by observing that the given sequence must have the form
{{ <c1+c2, left v1, C> |--> <c1,v1,LeftP c2 C> |-->* {[c1,v1', LeftP c2 C]} |--> {[c1+c2,left v1',C]} }}. 
For the strictly smaller inner subsequence by induction we have
{{c1~v1 |--> v1'}} which lets us complete the derivation of
{{c1+c2~(left v1) |--> left v1'}}. The {{right v2}} follows similarly.

@@

In other words, for any execution path from {{v}} to {{v'}} in {{c}},
{{c{dagger} }} has an execution path from {{v'}} to {{v}}.
Intuitively this maybe understood as follows: For every particular
execution sequence {{eta*}} transforms {{()}} into particular
{{(1/v,v)}}. On reverse execution the adjoint {{eps*}} will transform
{{(1/v,v)}} back to {{()}}.  Logical reversibility implies that
{{c^^^}} denotes the inverse relation that {{c}} denotes, i.e.
{{c^^^={|(v',v)~|~(v,v') elem c|} }}.

\roshan{This probably needs rigorous proof.}

@@

Like {{Pi}}, {{PiEE*}} is a commutative bimonoidal category.
Symmetric monoidal categories where every object had a dual object and
where dualizing preserves the monoidal tensor are referred to as
compact closed categories. Since division is defined only for types of
non-zero size, not all objects have duals.  We show that {{PiEE*}}
respects compact closed structure for non-zero objects. We introduce
the term ``locally compact closed'' to refer to this notion.
